---
title: "Practical ML Project"
author: "Cynthia"
date: "Wednesday, April 22, 2015"
output: html_document
---

#Practical Machine Learning - Course Project

##1. Preparation - load package, and set seed for reproduceability
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(1234)
```

##2. Getting and cleaning data - including deleting columns with all missing values and irrelavant columns
Note: Data for this project come from: http://groupware.les.inf.puc-rio.br/har.
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]
training<-training[,-c(1:7)]
testing <-testing[,-c(1:7)]
dim(training)
dim(testing)
```

##3. Cross-validation - partition training data by random sample without replacement of 75% and 25% rule
```{r}
subsamples <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
subTraining <- training[subsamples, ] 
subTesting <- training[-subsamples, ]
dim(subTraining)
dim(subTesting)
```
##4. Exploratory data analysis - graph the "class"" variable
```{r}
plot(subTraining$classe, col="lightgreen", main="Bar Plot of classe levels in subTraining dataset", xlab="classe levels", ylab="Frequency")
```

##5.1. Model building + test result by confusion matrix - decision tree
```{r}
model1 <- rpart(classe ~ ., data=subTraining, method="class")
prediction1 <- predict(model1, subTesting, type = "class")
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
confusionMatrix(prediction1, subTesting$classe)
```

##5.2. Model building + test result by confusion matrix - random forest
```{r}
model2 <- randomForest(classe ~. , data=subTraining, method="class")
prediction2 <- predict(model2, subTesting, type = "class")
confusionMatrix(prediction2, subTesting$classe)
```

##6. Conclusion and cross-validation
1. According to accuracy in the two tables(0.995 VS. 0.739), random forest performed better than decision tree, which is as expected. Hence, random forest model chosen. 
2. Expected out-of-sample error is estimated at 0.005, which is calculated as 1 - accuracy for predictions made against cross-validation. 
3. Test dataset comprises 20 cases. With an accuracy above 99% on cross-validation data, we expect very few test samples will be missclassified.
```{r}
predictfinal <- predict(model2, testing, type="class")
predictfinal
```